### 2019-01-22 macro3_1.tex
Macro

\begin{enumerate}
	\item Let $V(k)$ be the value of the agent who acts optimally and possesses $k$ units of capital now. Then Bellman equation is:
	\begin{align*}
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \beta V(k - c)\right\} \\
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \frac{1}{1+\rho} V(k - c)\right\} 
	\end{align*}
	Assume $c^*$ is the optimizer, then
	\begin{align*}
	V(k) = u(c^*) + \frac{1}{1+\rho}V(k-c^*)\\
	\rho = \frac{V(k - c^*) - V(k) + u(c^*)}{V(k) - u(c^*)}
	\end{align*}
	or alternatively:
	\begin{align*}
	\rho = \frac{u(c^*(t)) + V(k^*(t+1)) - V(k^*(t))}{V(k^*(t)) - u(c^*(t))}
	\end{align*}
	(* means optimal trajectory).
	Here $\frac{V(k(t+1)) - V(k(t))}{V(k(t))}$ are "capital gains" and $u(c^*)$ are "dividends".

@@

	\item Of course in general the answer to this question depends on the particular type of function $u()$, however if we assume "good" increasing and concave utility function then price of marginal piece of cake increases over time. It can be inferred both from the asset pricing equation and from the Euler's equation. From the asset pricing equation it is obviously that LHS is nonnegative (its meaning is the interest rate) hence the RHS should be also nonnegative. The denominator of the LHS is nonnegative because economically it is a shadow price of capital or matimatically it is equal to corresponding lagrange multiplier which cannot be negative by the Kuhn and Tucker theorem. Thus, numerator should be also nonnegative, so $V'(k(t+1)) \ge V'(k(t))$ i.e. price of marginal piece of cake increases over time. From the Euler's equation (which is obtained in 5.) $\frac{u'(c(t))}{u'(c(t+1))} = \frac{\lambda(t)}{\lambda(t+1)} = \beta$. Since $\lambda(t) = V'(k(t))$ then $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$. That means that $V'(k(t))$ is increasing over time. Intuitivelly it can be understood as follows: the amount of capital is shrinking hence over time the agent values each marginal unit of capital more and more. 
	\item Under some assumptions on the utility function it can be proved that $V(k)$ is strictly increasing. In our problem $k$ is decreasing over time that means that $V(k)$ should also decrease over time, i.e. the value of the whole cake decreases.

@@

intertemporal elasticity of substitution is $\frac{1}{\sigma}$. If it is equal to 0 then any changes of interest rate have no effect on consumption growth. If the intertemporal elasticity becomes larger than 0, then the ratio $\frac{c(t+1)}{c(t)}$ becomes low, that is the price of consumption 

>>>>

\begin{enumerate}
	\item Let $V(k)$ be the value of the agent who acts optimally and possesses $k$ units of capital now. Then Bellman equation is:
	\begin{align*}
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \beta V(k - c)\right\} \\
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \frac{1}{1+\rho} V(k - c)\right\} 
	\end{align*}
	Assume $c^*$ is the optimizer, then
	\begin{align*}
	V(k) = u(c^*) + \frac{1}{1+\rho}V(k-c^*)\\
	\rho = \frac{V(k - c^*) - V(k) + u(c^*)}{V(k) - u(c^*)}
	\end{align*}
	or alternatively:
	\begin{align*}
	\rho = \frac{u(c^*(t)) + V(k^*(t+1)) - V(k^*(t))}{V(k^*(t)) - u(c^*(t))}
	\end{align*}
	(* means optimal trajectory).
	Here $V(k(t+1)) - V(k(t))$ are gross "capital gains" and $u(c^*)$ are gross "dividends".

@@

	\item Of course in general the answer to this question depends on the particular type of function $u()$, however if we assume "good" increasing and concave utility function then price of marginal piece of cake increases over time. It can be inferred both from the asset pricing equation and from the Euler's equation. From the asset pricing equation it is obviously that LHS is nonnegative (its meaning is the interest rate) hence the RHS should be also nonnegative. The denominator of the RHS is nonnegative because economically it is a shadow price of capital or matimatically it is equal to corresponding lagrange multiplier which cannot be negative by the Kuhn and Tucker theorem. Thus, numerator should be also nonnegative, so $V'(k(t+1)) \ge V'(k(t))$ i.e. price of marginal piece of cake increases over time. From the Euler's equation (which is obtained in 5.) $\frac{u'(c(t))}{u'(c(t+1))} = \frac{\lambda(t)}{\lambda(t+1)} = \beta$. Since $\lambda(t) = V'(k(t))$ then $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$. That means that $V'(k(t))$ is increasing over time. Intuitivelly it can be understood as follows: the amount of capital is shrinking hence over time the agent values each marginal unit of capital more and more. 
	\item Under some assumptions on the utility function it can be proved that $V(k)$ is strictly increasing. In our problem $k$ is decreasing over time that means that $V(k)$ should also decrease over time, i.e. the value of the whole cake decreases. The amount of cake is decreasing over time, hence less cake $\to$ less it costs.

@@

intertemporal elasticity of substitution is $\frac{1}{\sigma}$. It shows the willingness of agents to substitute consumption over time i.e. to refuse today's consumption in order to increase it tomorrow. If $\frac{1}{\sigma} > 1$ then consumption path is steep: for the sake of increase consumtion today the agent is able to cut his consumption tomorrow, otherwise the consumption path is flat: the agent is not willing to cut tomorrow's consumption hence he should accept lower level of today's consumption (compare to the case $\frac{1}{\sigma} > 1$). See Fig. \ref{fig3}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig3}
\end{figure}
